---
title: "Week 1 - Homework"
author: "STAT 420, Summer 2017, Megan Masanz, netId: mjneuman"
date: "May 22, 2017"
output:
  html_document: 
    toc: yes
  pdf_document: default
---


## Exercise 1

For this exercise, we will use the `diabetes` dataset from the `faraway` package.

<!-- **Do** include the command to load the package into your environment. --> 
 
```{r}
library(faraway)
```


**(b)** Coerce the data to be a tibble instead of a data frame. 

<!--(You will need the `tibble` package to do so.) -->

```{r}
library(tibble)
diabetes = as_tibble(diabetes)
```

- Observations are in this dataset: **`r nrow(diabetes)`**

```{r}
nrow(diabetes)
```

- Number of variables in the dataset **`r ncol(diabetes)`**

```{r}
ncol(diabetes)
```


```{r}
str(diabetes)
```

- The dataset is composed of **403 African Americans in central Virginia**.
    + Ranging in age from 19-92
    + Number of Males: `r nrow(diabetes[ diabetes$gender == "male", ])` and Number of Females: `r nrow(diabetes[ diabetes$gender == "female", ])`

```{r}
range(diabetes$age)
nrow(diabetes[ diabetes$gender == "male", ])
nrow(diabetes[ diabetes$gender == "female", ])
```

**(c)** The following list provides the factor variables of the diabetes dataset

- location
- gender
- frame

```{r, echo = TRUE}
v = sapply(diabetes, is.factor, simplify = TRUE)
v[which(v == TRUE)]
```

**(d)** The mean [HDL](https://en.wikipedia.org/wiki/High-density_lipoprotein) level (High Density Lipoprotein) of individuals in this sample is **`r  mean(diabetes$hdl, na.rm = TRUE)`**

```{r}
mean(diabetes$hdl, na.rm = TRUE)
```

**(e)** The standard deviation of total cholesterol of individuals in this sample is **`r sd(diabetes$chol, na.rm = TRUE)`**

```{r}
sd(diabetes$chol, na.rm = TRUE)
```

**(f)** The range ages of individuals in this sample is from **`r  range(diabetes$age)[1]` **  to **`r  range(diabetes$age)[2]`**.

```{r}
range(diabetes$age)
```

**(g)** The mean HDL of females in this sample is **`r mean(diabetes[ diabetes$gender == "female", ]$hdl)`**

```{r}
mean(diabetes[ diabetes$gender == "female", ]$hdl)
```

**(h)** The Scatter plot of HDL (y-axis) vs weight (x-axis) found below shows there seems to be a relationship between the two variables.  The scatter plot appears to show a negative correlation (as weight increases, HDL decreases).

```{r, echo = TRUE}
plot(hdl ~ weight, data = diabetes,
     xlab = "Weight (in pounds)",
     ylab = "High Density Lipoprotein",
     main = "HDL vs Weight",
     pch  = 20,
     cex  = 1,
     col  = "dodgerblue")
```

**(i)** Based on the scatter plot of total cholesterol (y-axis) vs weight (x-axis) found below there does not appear to be a relationship between the two variables.  There does not seem to be a slope of the line of best fit for this data.

```{r, echo = TRUE}
plot(chol ~ weight, data = diabetes,
     xlab = "Weight (in pounds)",
     ylab = "Total Cholesterol",
     main = "Cholesterol vs Weight",
     pch  = 20,
     cex  = 1,
     col  = "dodgerblue")
```

**(j)** Below are the boxplots for HDL by gender. Based on the boxplot, there seems to be a difference between HDL levels for the 2 genders.  The median shows to be slightly higher for females, and based on the box plot the IQR is smaller for females than for males.  There also appears to be a higher number of interesting points/outliers in females than males.

```{r, echo = TRUE}
boxplot(hdl ~ gender, data = diabetes,
        xlab   = "Gender",
        ylab   = "High Density Lipoprotein",
        main   = "HDL vs Gender",
        pch    = 20,
        cex    = 1,
        col    = "darkorange",
        border = "dodgerblue")
```

## Exercise 2

For this exercise we will use the data stored in [`nutrition.csv`](nutrition.csv). 

```{r}
library(readr)
nutrition = read_csv("nutrition.csv")
```

It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA. It is a cleaned version totaling 5138 observations and is current as of September 2015.

The variables in the dataset are:

- `ID` 
- `Desc` - Short description of food
- `Water` - in grams
- `Calories` - in kcal
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - Carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - Vitamin C, in milligrams
- `Chol` - Cholesterol, in milligrams
- `Portion` - Description of standard serving size used in analysis

**(a)** Below is a histogram of `Calories` with `R`'s default bin selection. The histogram is showing a right skewed shape with an interesting increase in values between 350-400.

```{r, echo = TRUE}
hist(nutrition$Calories,
     xlab   = "Calories",
     main   = "Histogram of Calories",
     col    = "dodgerblue",
     border = "darkorange")
```

**(b)** Below is a scatter plot of calories (y-axis) vs protein (x-axis). 
One trend is that the standard deviation for Calories seems to decrease as protein increases.  Based on the high variability, a good prediction could not be made on the calories in food based on protein alone.

```{r, echo = TRUE}
plot(Calories ~ Protein, data = nutrition,
     xlab = "Protein",
     ylab = "Calories",
     main = "Calories vs Protein",
     pch  = 20,
     cex  = 1,
     col  = "dodgerblue")
```

**(c)** Below is a scatter plot of `Calories` (y-axis) vs `4 * Protein + 4 * Carbs + 9 * Fat + 2 * Fiber` (x-axis). 

The results of this equation provide a straight line with some variability which could be due to a variety of factors.  As an example in a given serving size the calories could be within a range of values.

```{r, echo = TRUE}
calorieCalc = 4 * nutrition$Protein + 4 * nutrition$Carbs + 9 * nutrition$Fat + 2 * nutrition$Fiber
nutrition = cbind(nutrition, calorieCalc)

plot(Calories ~ calorieCalc, data = nutrition,
     xlab = "4 * Protein + 4 * Carbs + 9 * Fat + 2 * Fiber",
     ylab = "Calories",
     main = "Calories vs \n Calorie count based on the protein, carbohydrate, and fat values",
     pch  = 20,
     cex  = 1,
     col  = "dodgerblue")

```

## Exercise 3

For each of the following parts, use the following vectors:

```{r}
a = 1:10
b = 10:1
c = rep(1, times = 10)
d = 2 ^ (1:10)
```

**(a)** Function called `sum_of_squares`.

- Arguments:
    - A vector of numeric data `x`.
- Output:
    - The sum of the squares of the elements of the vector. $\sum_{i = 1}^n x_i^2$
    
Provide your function, as well as the result of running the following code:
```{r function_sum_of_squares}
sum_of_squares = function(x){
  sum(x ^ 2)
}
```

```{r, eval = TRUE}
sum_of_squares(x = a)
sum_of_squares(x = c(c, d))
```

**(b)** Write a function called `rms_diff`.

- Arguments:
    - A vector of numeric data `x`.
    - A vector of numeric data `y`.
- Output:
    - $\sqrt{\frac{1}{n}\sum_{i = 1}^n (x_i - y_i)^{2}}$

Provide your function, as well as the result of running the following code:

```{r}
rms_diff = function (x, y){
    sqrt(  (1/length(x)) * sum(( x - y) ^ 2))
}
```

```{r, eval = TRUE}
rms_diff(x = a, y = b)
rms_diff(x = d, y = c)
rms_diff(x = d, y = 1)
rms_diff(x = a, y = 0) ^ 2 * length(a)
```

