---
title: "Final Project Analysis"
author: "Megan"
date: "July 26, 2017"
output: html_document
---

```{r}
library(MASS)
library(readr)
library(lmtest)
library(car)
library(leaps)
library(knitr)
library(faraway)
library(tibble)


fb = read.table("dataset_Facebook.csv", sep = ";", header = T)



fb$Type           = as.factor(fb$Type)
fb$Category       = as.factor(fb$Category)
fb$Post.Month     = as.factor(fb$Post.Month)
fb$Post.Hour      = as.factor(fb$Post.Hour)
fb$Post.Weekday   = as.factor(fb$Post.Weekday)
fb$Paid           = as.factor(fb$Paid)






full_model = lm(Lifetime.Post.Consumers ~ ., data = fb)

colnames(fb)

typeLevels = levels(fb$Type)

summary(full_model)
fb$Total.Interactions = NULL
full_model = lm(Lifetime.Post.Consumers ~ ., data = fb)
summary(full_model)
nrow(fb)
#Cleansing Data
fb = fb[complete.cases(fb), ]

#Initial Analysis:
variables = colnames(fb)
predictor_variables = variables[which(variables != "Lifetime.Post.Consumers")]
response_variable = "Lifetime.Post.Consumers"



getAdditiveModels = function(variables, response_variable, data_set, alpha)
{
  num_predictor_variables = length(variables)
  p_val           = rep(0, num_predictor_variables ^ 2)
  bp_p_value      = rep(0, num_predictor_variables ^ 2)
  shapiro_p_value = rep(0, num_predictor_variables ^ 2)
  strmodel        = rep(0, num_predictor_variables ^ 2)
  rss             = rep(0, num_predictor_variables ^ 2)
  beta_parameter  = rep(0, num_predictor_variables ^ 2)
  pass_bp         = rep(0, num_predictor_variables ^ 2)
  pass_shapiro    = rep(0, num_predictor_variables ^ 2)
  adjustedR2      = rep(0, num_predictor_variables ^ 2)
  sizeResiduals   = rep(0, num_predictor_variables ^ 2)
  
  index = 1;
  
  for (j in 1:num_predictor_variables)
  {
    smallestrss = 0
    smallestrssIndex = 0 
    for (i in 1:length(variables))
    {
      if(j == 1) #at the beginning
      {
        startpointer          = index
        #could make s1 be log(y)
        s1                    = response_variable
        s2                    = variables[i]
        strmodel[index]       = paste(s1, " ~ " , s2)
        
      }
      else
      {
        s2                    = variables[i]
        strmodel[index]       = paste(previousBestModel , " + " , s2)
        
      }
      
      model                    = lm(strmodel[index], data = data_set)
      rss[index]               = sum(resid(model)^2) 
      bp_p_value[index]        = bptest(model)$p.value
      shapiro_p_value[index]   = shapiro.test(resid(model))$p.value
      beta_parameter[index]    = j
      adjustedR2[index]        = summary(model)$adj.r.squared
      sizeResiduals[index]     = length(resid(model))
      
      if (sizeResiduals[i] == 0 )
      {
        print("ERROR!!!!!!!!!!!!!!!!!!!")
        
      }
      else
        print(paste(strmodel[index], ": " , sizeResiduals[i]))
              
      
      if (shapiro_p_value[i] > alpha) {
        pass_shapiro[index] = TRUE
      }
      
      if (bp_p_value[i] > alpha) {
        pass_bp[index] = TRUE
      }
      
      if (smallestrssIndex == 0 || rss[index] < smallestrss){
        
        smallestrssIndex = index
        smallestrss = rss[index]
      }
      index = index + 1
      
    }
    #for a given number of predictor variables create temp vector and find the min rss
    previousBestModel = strmodel[smallestrssIndex]
  }
  result         = data.frame(beta_parameter, strmodel, rss, adjustedR2, bp_p_value, shapiro_p_value, pass_bp, pass_shapiro, sizeResiduals)
  result$strmodel =  as.character(result$strmodel)
  result
}


results = getAdditiveModels(predictor_variables, response_variable, fb, .05)


#here is a function that for a given number of beta parameters pulls out the lowest rss
getModels = function(results, num_predictor_variables){
  bestModelForParamNumber = rep("", num_predictor_variables)
  for (i in 1:num_predictor_variables)
  {
    subset_data = results[results$beta_parameter == i, ]
    
    bestModelForParamNumber[i] = subset_data$strmodel[which.min(subset_data$rss)]
    
  }
  bestModelForParamNumber
}


bestModels = getModels(results, length(predictor_variables))


#now for each of the "best" linear models, do an anova test to determine which are actually the best
doAnova = function(dataset, models, alpha)
{
  p_val   = rep(0, length(models) - 1)
  keep_old = rep(FALSE, length(models) - 1)
  winning_model = rep(0, length(models) - 1)
  stop = FALSE
  #print(length(models))
  for (i in 1:(length(models) -1))
  {
    if (stop == FALSE)
    {
      p_val[i] = anova(lm(bestModels[i], data = dataset), lm(bestModels[i + 1], data = dataset))$"Pr(>F)"[2]
      if (p_val[i] < alpha)
      {
        keep_old[i] = FALSE
        winning_model[i] = bestModels[i + 1]
        stop = FALSE
      }
      else
      {
        keep_old[i] = TRUE
        winning_model[i] = bestModels[i]
        stop = TRUE
        report_winning  = winning_model[1:i]
        report_p_val    = p_val[1:i]
        report_keep_old = keep_old[1:i]
      }
    }
  }
  
  output = data.frame(report_winning, report_p_val, report_keep_old)
  output$report_winning =  as.character(output$report_winning)
  output
}


results = doAnova(fb, bestModels, 0.05 )

model_data_selected_RSS_Manual = results[results$report_keep_old == TRUE, ]

model_selected_RSS_Manual = lm(model_data_selected_RSS_Manual$report_winning, data = fb)


#best additive model based on RSS and Anova Test
#Lifetime.Post.Consumers  ~  Lifetime.Engaged.Users  +  like  +  share  +  Lifetime.People.who.have.liked.your.Page.and.engaged.with.your.post  +  Post.Month  +  Lifetime.Post.Consumptions  +  comment  +  Lifetime.Post.Total.Reach

###########################################################
#AIC and BIC
##########################################################
predictor_variables = variables[which(variables != "Lifetime.Post.Consumers")]
length(predictor_variables)

createStringFullModel = function(predictor_variables){
  num_predictor_variables = length(predictor_variables)
  for (i in 1:num_predictor_variables)
  {
    if (i == 1)
    {
      model = paste("Lifetime.Post.Consumers", "~" ,predictor_variables[i] )
      print(model)
    }
    else
      model = paste(model , " + " , predictor_variables[i])
  }
  model
}

strFullModel = createStringFullModel(predictor_variables)


##########################################################
#AIC - less aggressive
##########################################################
#forward AIC
fb_mod_start = lm(Lifetime.Post.Consumers ~ 1, data = fb)
model_forward_aic = step(fb_mod_start,
                         strFullModel,
                         direction = "forward", trace = 0)

summary(model_forward_aic)

vif(model_forward_aic)

fullMOdel = lm(strFullModel, data = fb)
model_back_aic = step(fullMOdel, direction = "backward", trace = 0)
summary(model_back_aic)
vif(model_back_aic)



##########################################################
#BIC - more aggressive
##########################################################
full_model = lm(Lifetime.Post.Consumers ~ ., data = fb)
n = length(resid(full_model))
model_forward_bic = step(fb_mod_start,
                         strFullModel,
                         direction = "forward", k = log(n), trace = 0)
summary(model_forward_bic)

vif(model_forward_bic)

#backward BIC
model_back_bic = step(full_model, direction = "backward", k = log(n), trace = 0)
summary(model_back_bic)

###########################################################
#Make a table saying which variables are in which model
#to specify which variables to explore
##########################################################



                                                         
all = predictor_variables %in% names(coef(model_selected_RSS_Manual)) %in%  names(coef(model_forward_aic)) %in% names(coef(model_back_aic)) %in% names(coef(model_forward_bic)) 

Variables_RSS_Manual   = predictor_variables %in% names(coef(model_selected_RSS_Manual))
Variables_Forward_AIC  = predictor_variables %in% names(coef(model_forward_aic))
Variables_Backward_AIC = predictor_variables %in% names(coef(model_back_aic))
Variables_Forward_BIC  = predictor_variables %in% names(coef(model_forward_bic))
Variables_Backward_BIC = predictor_variables %in% names(coef(model_back_bic))

#Variable, RSS Evelaution per beta parameter, Backward BIC, Forward BIC, Backward AIC, Forward AIC
variables_for_analysis = data.frame(predictor_variables, Variables_RSS_Manual, Variables_Forward_BIC, Variables_Backward_BIC, Variables_Forward_AIC, Variables_Backward_AIC)

colnames(variables_for_analysis) = c("Variables" , "Variable included in RSS Manual", "Variable included in Forward BIC", "Variable included in Backward BIC", "Variable included in Forward AIC", "Variable included in Backward AIC")
knitr::kable(variables_for_analysis, caption = "MLR Additive Model Variable Analysis")
```

This table gives us a good idea of what variables we should use in our analysis.  We can now begin with looking at a pairs plot with the subset of variables that are included in all of the additive models to better understand the predictor variables themselves.


```{r}
fb_subset = subset(fb, select = c("Lifetime.Post.Consumers", "Lifetime.Post.Total.Reach", "Lifetime.Engaged.Users", "Lifetime.Post.Consumptions", "Lifetime.People.who.have.liked.your.Page.and.engaged.with.your.post", "comment", "like", "share"))

```

The pairs plot for this subset of variables is found below
```{r fig.height=20, fig.width=10}
pairs(fb_subset)
```

With the pairs plot, we can see it appears some relationships could be looked at, but quickly seeing the order of magnitude on the reponse variable, suggests we should also do a box-cox plot to determine if perhaps the response should be transformed.

```{r}
boxcox(model_selected_RSS_Manual, plotit = TRUE)

par(mfrow = c(2,2))
boxcox(model_forward_aic, plotit = TRUE)
boxcox(model_back_aic, plotit = TRUE)
boxcox(model_forward_bic, plotit = TRUE)
boxcox(model_back_bic, plotit = TRUE)

```

The box-cox plot for all the models currently created suggests a transformation of the response variable

```{r fig.height=20, fig.width=10}
fb_subset = subset(fb, select = c("Lifetime.Post.Consumers", "Lifetime.Post.Total.Reach", "Lifetime.Engaged.Users", "Lifetime.Post.Consumptions", "Lifetime.People.who.have.liked.your.Page.and.engaged.with.your.post", "comment", "like", "share"))

log.Consumers = log(fb_subset$Lifetime.Post.Consumers)

fb_withResponseTransform = cbind( fb_subset, log.Consumers)
pairs(fb_withResponseTransform, col = "dodgerblue")

```

The pairs plot shown above makes seeing a few trends much easier with the addition of the log.Consumers.  Several variables appear to either require a log transformation, or perhaps a power transforation.  Given we currently have 7 predictors of interest, a function was written to automate reviewing this information.

Generically we will add + 1 to each predictor variable for log transformation to ensure the values are great than 0 if the min val

```{r}
min(fb_withResponseTransform$Lifetime.Post.Total.Reach)
min(fb_withResponseTransform$Lifetime.Engaged.Users)
min(fb_withResponseTransform$Lifetime.Post.Consumptions)
min(fb_withResponseTransform$Lifetime.People.who.have.liked.your.Page.and.engaged.with.your.post)
min(fb_withResponseTransform$comment)
min(fb_withResponseTransform$like)
min(fb_withResponseTransform$share)

```


```{r}

transformationsPowers = function(predictor, response, dataset, responsevector, predictorvector){
  
  strQuad                      = paste("log(", response, ") ~ ", predictor , " + I(" , predictor, "^ 2) ")
  
  quad_model                   = lm(strQuad, data = dataset)
  
  quad_RMSE                    = sqrt(mean((responsevector - exp(fitted(quad_model))) ^ 2))
  quad_bp_p_value              = bptest(quad_model)$p.value
  quad_shapiro_p_value         = shapiro.test(resid(quad_model))$p.value
  
#  quad = c(quad_RMSE, quad_bp_p_value, quad_shapiro_p_value)
  
  strThird                     = paste("log(", response, ") ~ ", predictor , " + I(" , predictor, "^ 2)  + I(", predictor, "^3)")
  third_model                  = lm(strThird, data = dataset)
  
  third_RMSE                   = sqrt(mean((responsevector - exp(fitted(third_model))) ^ 2))
  third_bp_p_value             = bptest(third_model)$p.value
  third_shapiro_p_value        = shapiro.test(resid(third_model))$p.value
  
#  third = c(third_RMSE, third_bp_p_value, third_shapiro_p_value)
  
#  result = data.frame(log = log, quad = quad, third = third)
 
  predictor   = rep(predictor, 2)
  transform   = c("^2 obeying hierachy", "^3 obeying hierachy")
  bp          = c(quad_bp_p_value, third_bp_p_value)
  shapiro     = c(quad_shapiro_p_value, third_shapiro_p_value)
  RMSE        = c(quad_RMSE, third_RMSE)
  
  data.frame(predictor = predictor, transform = transform, bp = bp, shapiro = shapiro, RMSE = RMSE)
  
}

variables = colnames(fb_withResponseTransform)
predictor_variables = variables[which((variables != "Lifetime.Post.Consumers") & variables != "log.Consumers")]

length(predictor_variables)

result1 = transformationsPowers(predictor_variables[1], "Lifetime.Post.Consumers", fb_withResponseTransform, fb_withResponseTransform$Lifetime.Post.Consumers)
knitr::kable(result1, caption = "Transformation Review")

result1 = transformationsPowers(predictor_variables[2], "Lifetime.Post.Consumers", fb_withResponseTransform, fb_withResponseTransform$Lifetime.Post.Consumers)
knitr::kable(result1, caption = "Transformation Review")

result1 = transformationsPowers(predictor_variables[3], "Lifetime.Post.Consumers", fb_withResponseTransform, fb_withResponseTransform$Lifetime.Post.Consumers)
knitr::kable(result1, caption = "Transformation Review")

result1 = transformationsPowers(predictor_variables[4], "Lifetime.Post.Consumers", fb_withResponseTransform, fb_withResponseTransform$Lifetime.Post.Consumers)
knitr::kable(result1, caption = "Transformation Review")

result1 = transformationsPowers(predictor_variables[5], "Lifetime.Post.Consumers", fb_withResponseTransform, fb_withResponseTransform$Lifetime.Post.Consumers)
knitr::kable(result1, caption = "Transformation Review")

result1 = transformationsPowers(predictor_variables[6], "Lifetime.Post.Consumers", fb_withResponseTransform, fb_withResponseTransform$Lifetime.Post.Consumers)
knitr::kable(result1, caption = "Transformation Review")

result1 = transformationsPowers(predictor_variables[7], "Lifetime.Post.Consumers", fb_withResponseTransform, fb_withResponseTransform$Lifetime.Post.Consumers)
knitr::kable(result1, caption = "Transformation Review")
```
Next we will examine log transformations


```{r}



transformationsLog = function (response, predictors, dataset, responsevector){
  log_RMSE = rep(0:length(predictors))
  log_bp_p_value  = rep(0:length(predictors))
  log_shapiro_p_value  = rep(0:length(predictors))
  
for (i in 1: length(predictors)){

  print( predictors[i] 
         )
  strLog = paste("log(", response, ") ~ log(", predictors[i] , ")")

  log_model = lm(strLog, data = dataset)
  
  log_RMSE[i]                     = sqrt(mean((responsevector - exp(fitted(log_model))) ^ 2))
  log_bp_p_value[i]               = bptest(log_model)$p.value
  log_shapiro_p_value[i]          = shapiro.test(resid(log_model))$p.value
}
  
  transform = rep("log of predictor", length(predictors))
  data.frame( transform = transform, bp = log_bp_p_value, shapiro = log_shapiro_p_value, RMSE = log_RMSE )
  
}
```

In order to explore log transformations on predictor variables, in the case when the range is from 0 to some value, we will add 1 to the predictor variable.
```{r}
fb_withResponseTransformForLog = fb_withResponseTransform
fb_withResponseTransformForLog$like = fb_withResponseTransformForLog$like + 1
fb_withResponseTransformForLog$share = fb_withResponseTransformForLog$share + 1
fb_withResponseTransformForLog$comment = fb_withResponseTransformForLog$comment + 1

(predictor_variables)

result1 = transformationsLog("Lifetime.Post.Consumers", predictor_variables, fb_withResponseTransformForLog, fb_withResponseTransformForLog$Lifetime.Post.Consumers )

knitr::kable(result1, caption = "Transformation Review")
```


predictor = predictors, transform = transform,



