---
title: "Final Project"
author: "Megan"
date: "July 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(readr)
library(lmtest)
library(car)
library(leaps)
library(knitr)
fb <- read_delim("dataset_Facebook.csv", ";", escape_double = FALSE, trim_ws = TRUE)
```

Before analyzing the data, we will explore what the variables mean, and determine which variables should be considered as Factor Variables.

Variables:

- `Type` : `Link`, `Photo`, `Status`, `Video`
- `Category`: `1`, `2`, `3`
- `Post Month`

```{r}
fb$Type           = as.factor(fb$Type)
fb$Category       = as.factor(fb$Category)
fb$`Post Month`   = as.factor(fb$`Post Month`)
fb$`Post Hour`    = as.factor(fb$`Post Hour`)
fb$`Post Weekday` = as.factor(fb$`Post Weekday`)
fb$Paid           = as.factor(fb$Paid)



result = cbind( "Type",         levels(fb$Type), 
                "Category",     levels(fb$Category),
                "Post Month",   levels(fb$`Post Month`),
                "Post Hour",    levels(fb$`Post Hour`),
                "Post Weekday", levels(fb$`Post Weekday`),
                "Paid",         levels(fb$Paid)
                )


kable(result, format = "markdown", padding = 3)
```



An effort review the variance inflation factors in R causes a problem since there are aliased coeffients in the model.  First we will need to explore the data to understand where there are aliased coefficents.


