---
title: "Quiz 2"
author: "Megan Masanz"
date: "May 26, 2017"
output: html_document
---

Currently incorrect: 13


1.  (correct) Consider the linear regression model

$$
Y = 10 + 5x + \epsilon
$$

$$
\epsilon \sim N(0, \sigma^2 = 16)
$$
Calculate the probability that Y is less than 6 given that x=0.

Answer:  0.1586553

```{r}
pnorm(-4, 0 ,4 ) 
```

2. (correct)  using the SLR model in question 1, what is the probability that Y is greater than 3 given that x = -1?


```{r}
pnorm(-2,0,4, lower.tail = FALSE)
```


3.(correct) Using the SLP model in Question 1, what is the prob that Y is greater than 3
given that X = -2?

```{r}
pnorm(3, 0, 4, lower.tail = FALSE)
```


4.  (correct) Fit a SLR to predict duration of an eruption of the old faithful based on waiting time

What is the estimate of the intercept parameter?

```{r}
waiting_model = lm(eruptions ~ waiting, data = faithful)
summary(waiting_model)$coefficients[1]
```

5. (correct)  What is the estimate of the slope parameter
```{r}
waiting_model = lm(eruptions ~ waiting, data = faithful)
coef(waiting_model)[2]
```

6.  (correct) Predict duration for time of 80 minutes

```{r}
waiting_model = lm(eruptions ~ waiting, data = faithful)
predict(waiting_model, newdata = data.frame(waiting = 80))
```

7. (correct) Predict for 120 minutes
```{r}
waiting_model = lm(eruptions ~ waiting, data = faithful)
predict(waiting_model, newdata = data.frame(waiting = 120))
```


8.  (correct) Which is more accurate - the 80 since 80 is in the dataframe
```{r}
80 %in% faithful$waiting
```

9.  **Incorrect?** Calculate RSS for the model
```{r}
waiting_model = lm(eruptions ~ waiting, data = faithful)
# get yi_hat values for the faithful dataset
yi_hat = predict(waiting_model, newdata = data.frame(faithful))
##RSS = Residual Sum of Squares
(sum( (faithful$eruptions  - yi_hat )^ 2))/ (length(faithful$eruptions) - 2)

## Alternatively
sum(waiting_model$residuals^2)/(length(faithful$eruptions) - 2)
```

9 - rethinking question 9 - better not be this answer
```{r}
sum(waiting_model$residuals ^ 2)
```

10.  What proportion of the variation in eruption duration is explained by the linear relationship with waiting time?
```{r}
waiting_model = lm(eruptions ~ waiting, data = faithful)
summary(waiting_model)$r.squared
```

11.  Calculate the standard deviation of the residuals of the fitted model.

```{r}
waiting_model = lm(eruptions ~ waiting, data = faithful)
sd(waiting_model$residuals)
```

**But questioning since I thought it was**
```{r}
#Question 11
summary(waiting_model)$sigma

#Question 11 also answered
waiting_model = lm(eruptions ~ waiting, data = faithful)
fitted_data = predict(waiting_model, newdata = data.frame(faithful))
sqrt((sum( (faithful$eruptions  - fitted_data )^ 2))/ (length(fitted_data) - 2))
```


12.  Suppose both least squares and maximum likelihood are used to fit a simple linear regression model to the same data. The estimates for the slope and the intercept will be:

The same.

13. ** None are true. -  I got this wrong. **

Consider the fitted regression model:

y^= ???1.5+2.3x
Select all of the following that must be true.

The difference between the y values of observations at x=10 and x=9 is 2.3.
(no)

There are observations in the dataset used to fit this regression with negative y values.
(no)

A good estimate for the mean of Y when x=0 is -1.5.
(no - i guess if i had to pic an option?)

14. Selected last 2 options.
The SLR model assumes that the relationship between the response and the predictor is linear. 
The SLR model assumes that errors are independent. (last option)


15. Suppose you fit a simple linear regression model and obtain ??^=0.

Does this mean that there is no relationship between the response and the predictor?

Answer: No

